Packaging research artefacts with RO-Crate

### 2022-01-02 latex/ro-crate.tex
Small typos

It is important to note that having all these computational details in
order to represent them in an RO-Crate is an ideal scenario -- in
practice there will always be gaps of knowledge, and exposing all
provenance details automatically would require improvements to the data
sources, workflow, workflow engine and its dependencies. RO-Crate can
be seen as a flexible annotation mechanism for augmenting automatic
workflow provenance. Additional metadata can be added manually, e.g.
for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{M etadata are accessible, even when the
data are no longer available} \cite{doi:10.1038/sdata.2016.18}.}, or
to cite software that lack persistent identifiers. This inline
\textit{FAIRifying} allows researchers to achieve ``just enough FAIR'' to
explain their computational experiments.

@@

RO-Crate is formed by a set of best practice recommendations, developed
by an open and broad community. These guidelines show how to use ``just
enough'' Linked Data standards in a consistent way. The use of
structured metadata with a rich base vocabulary can cover
general-purpose contextual relations, with a Linked Data foundation
that ensures extensibility to domain- and application-specific uses. We
can therefore consider an RO-Crate not just as a structured data
archive, but as a multimodal scholarly knowledge graph that can help
``FAIRify'' and combine metadata of existing resources.

The adoption of simple Web technologies in the RO-Crate specification
has helped a rapid development of a wide variety of supporting open
source tools and libraries. RO-Crate fits into the larger landscape of
open scholarly communication and FAIR Digital Object infrastructure,
and can be integrated into data repository platforms. RO-Crate can be
applied as a data/metadata exchange mechanism, assist in long-term
archival preservation of metadata and data, or simply used at a small
scale by individual researchers. Thanks to its strong community
support, new and improved profiles and tools are being continuously
added to the RO-Crate tooling landscape, making it easier for adopters
to find examples and support for their own use case.

@@

We operate in an open community where future and potential users of
RO-Crate are actively welcomed to participate and contribute feedback
and requirements. In addition we are targeting a wider audience through
extensive outreach
activities\footnote{\url{https://www.researchobject.org/ro-crate/outreach.html}} and
by initiating new connections. Recent contacts include American
Geophysical Union (AGU) on Data Citation Reliquary
\cite{doi:10.5281/zenodo.4916734}, National Institute of Standards and
Technology (NIST) on material science and
InvenioRDM\footnote{\url{https://inveniosoftware.org/products/rdm/}} used by the
Zenodo data repository. New Horizon Europe projects adapting RO-Crate
include BY-COVID,\footnote{\url{https://by-covid.org/}} which aims to improve FAIR
access to data on COVID-19 and other infectious diseases.

The main addition in the upcoming 1.2 release of the RO-Crate
specifications will be the formalization of
profiles\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
for different categories of crates. Additional entity types have been
requested by users, e.g. workflow runs, business workflows, containers
and software packages, tabular data structures; these are not always
matched well with existing Schema.org types but may benefit from other
vocabularies or even separate metadata files, e.g. from Frictionless
Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning with
and collaborating with related research artefact description efforts
like CodeMeta\footnote{\url{https://codemeta.github.io/}} for software metadata,
Science-on-Schema.org\footnote{\url{https://science-on-schema.org/}}
\cite{doi:10.5281/zenodo.4477164} for datasets, FAIR Digital
Objects\footnote{\url{https://fairdo.org/}} \cite{doi:10.3390/publications8020021} and
activities in EOSC task forces\footnote{\url{https://www.eosc.eu/task-force-faq}}
including the EOSC Interoperability Framework \cite{doi:10.2777/620649}.

@@

Author contributions to this article and the RO-Crate projet according
to the Contributor Roles Taxonomy CASRAI
CrEDiT\footnote{\url{https://casrai.org/credit/}} \cite{doi:10.1087/20150211}:
\begin{description}
\item[Stian Soiland-Reyes]  Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project administration,
Software, Visualization, Writing -- original draft, Writing -- review
\& editing

>>>>

It is important to note that having all these computational details in
order to represent them in an RO-Crate is an ideal scenario -- in
practice there will always be gaps of knowledge, and exposing all
provenance details automatically would require improvements to the data
sources, workflow, workflow engine and its dependencies. RO-Crate can
be seen as a flexible annotation mechanism for augmenting automatic
workflow provenance. Additional metadata can be added manually, e.g.
for sensitive clinical data that cannot be publicly exposed\footnote{FAIR principle A2: \textit{Metadata are accessible, even when the
data are no longer available} \cite{doi:10.1038/sdata.2016.18}.}, or
to cite software that lack persistent identifiers. This inline
\textit{FAIRifying} allows researchers to achieve ``just enough FAIR'' to
explain their computational experiments.

@@

RO-Crate is formed by a set of best practice recommendations, developed
by an open and broad community. These guidelines show how to use ``just
enough'' standards in a consistent way. The use of
structured metadata with a rich base vocabulary can cover
general-purpose contextual relations, with a Linked Data foundation
that ensures extensibility to domain- and application-specific uses. We
can therefore consider an RO-Crate not just as a structured data
archive, but as a multimodal scholarly knowledge graph that can help
``FAIRify'' and combine metadata of existing resources.

The adoption of simple Web technologies in the RO-Crate specification
has helped a rapid development of a wide variety of supporting open
source tools and libraries. RO-Crate fits into the larger landscape of
open scholarly communication and FAIR Digital Object infrastructure,
and can be integrated into data repository platforms. RO-Crate can be
applied as a data/metadata exchange mechanism, assist in long-term
archival preservation of metadata and data, or simply used at a small
scale by individual researchers. Thanks to its strong community
support, new and improved profiles and tools are being continuously
added to the RO-Crate landscape, making it easier for adopters
to find examples and support for their own use case.

@@

We operate in an open community where future and potential users of
RO-Crate are actively welcomed to participate and contribute feedback
and requirements. In addition, we are targeting a wider audience through
extensive outreach
activities\footnote{\url{https://www.researchobject.org/ro-crate/outreach.html}} and
by initiating new connections. Recent contacts include American
Geophysical Union (AGU) on Data Citation Reliquary
\cite{doi:10.5281/zenodo.4916734}, National Institute of Standards and
Technology (NIST) on material science, and
InvenioRDM\footnote{\url{https://inveniosoftware.org/products/rdm/}} used by the
Zenodo data repository. New Horizon Europe projects adapting RO-Crate
include BY-COVID,\footnote{\url{https://by-covid.org/}} which aims to improve FAIR
access to data on COVID-19 and other infectious diseases.

The main addition in the upcoming 1.2 release of the RO-Crate
specifications will be the formalization of
profiles\footnote{\url{https://www.researchobject.org/ro-crate/1.2-DRAFT/profiles}}
for different categories of crates. Additional entity types have been
requested by users, e.g. workflow runs, business workflows, containers
and software packages, tabular data structures; these are not always
matched well with existing Schema.org types, but may benefit from other
vocabularies or even separate metadata files, e.g. from Frictionless
Data.\footnote{\url{https://frictionlessdata.io/}} We will be further aligning
and collaborating with related research artefact description efforts
like CodeMeta\footnote{\url{https://codemeta.github.io/}} for software metadata,
Science-on-Schema.org\footnote{\url{https://science-on-schema.org/}}
\cite{doi:10.5281/zenodo.4477164} for datasets, FAIR Digital
Objects\footnote{\url{https://fairdo.org/}} \cite{doi:10.3390/publications8020021} and
activities in EOSC task forces\footnote{\url{https://www.eosc.eu/task-force-faq}}
including the EOSC Interoperability Framework \cite{doi:10.2777/620649}.

@@

Author contributions to this article and the RO-Crate project according
to the Contributor Roles Taxonomy CASRAI
CrEDiT\footnote{\url{https://casrai.org/credit/}} \cite{doi:10.1087/20150211}:
\begin{description}
\item[Stian Soiland-Reyes]  Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project administration,
Software, Visualization, Writing -- original draft, Writing -- review
\& editing
